Order reduction and partial MILP models \\ for certifying Robustness of DNNs globally.

### 2025-07-29 AAAI2026/Lipitschz.tex
global

	\section{Experimental Evaluation}

	We implemented our code in Python 3.8.
	Gurobi 9.52 was used for solving LP and MILP problems. We conducted our evaluation on an AMD Threadripper 7970X  ($32$ cores$@4.0$GHz) with 256 GB of main memory and 2 NVIDIA RTX 4090.

	We will do experiments on two networks for MNIST dataset and another physical model. Two MNIST have the same structure. They are full-connected DNN with 5, 10 outputs and 784 inputs; each hidden layer has 100 neurons. One of them is a normally trained neural network and another is for trained for adversarial robustness. We call them MNIST $5\times100$-Normal and MNIST $5\times 100$-DiffAI. As the name shows MNIST $5\times 100$-DiffAI has a better robustness. The physical network is a simplified model with 10 input neurons, 26 output neurons, and 2 hidden layers with 50 neurons each.

	\subsection{Experiment results for robustness (MNIST)}

	\begin{table}[h!]
	\begin{tabular}{|l|c|c|c|c|}\hline\hline
		$L_1\leq 0.5$ &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ &  Time \\\hline \hline
		1v, $375 \times 1$ & 17.4724 & 0.9089  ?? & 0.0917 & 43200s \\\hline 
		2v, $375 \times 2$ & 21.0135 & 8.4672& 0.2229 & 43200s \\\hline		
		2v, $450 \times 2$ & 12.0097 & 3.4408 & {\bf 0.2584} & 43200s \\\hline
		1v, $500 \times 1$ & ?? & ?? & ?? & 43200s \\\hline 
		3v, $500 \times 3$ & ?? & ?? & ?? & 43200s \\\hline 
	 2v, $500 \times 2$ & {\bf 8.7608} & nan & nan & 43200s \\\hline\hline

	 2v, $485 \times 2$ & 20.6460 & 1.9152 & {\bf 0.3722} & 130000s \\\hline\hline

		 2v, $500 \times 2$ & {\bf 8.2573} & nan & nan & 260000s \\\hline\hline

	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models 
	to obtain bounds on $\beta^{0.5}_{0,1}$ on the {\bf full dimension} MNIST DNN, for timeouts of 43200s, 130000s and 260000s, where 375, 450, 485 or 500 ($\times 1$, $\times 2$, $\times 3$) neurons use binary variables.}
\end{table}

On the full space, we reach bounds of $8.2$ (on $\beta^{.5}_{0,1}$), which is too pessimistic as it allows to certify 0 image robust. The worst-case found by the "2v" $485 \times 2$ binary variables after 130000s only displays a difference of 
$0.37$, $20 \times$ away from the bound found.

\begin{figure*}[t!]
	\centering
\includegraphics[scale=0.5]{image.png} \hspace{0.8cm}
\includegraphics[scale=0.5]{perturb.png}
\caption{An improbable image and its perturbation (difference around x=11, y=18) 
with maximal $\beta^{.5}_{0,1}=0.37$ for MNIST as obtained by the "2v" $485 \times 2$ model in full 758 dimension image space.}
\end{figure*}

\paragraph{Reduced space}

We now consider a PCA model order reduction to avoid considering improbable images, 
speed up obtaining the bounds, and obtain less pessimistic bounds.
We settle on 20 orders, as the MNIST DNN considered run on a images obtained from projecting to the reduced order and projected back to the full dimension 
display the same accuracy of $97$\% as the DNN on the original images, meaning no accuracy is lost from reducing to 20 dimensions, which is the only thing which matters. On such a reduced space, bounds obtained are much more precise, and one can certify robustness of images online.

	\begin{table}[h!]
	\begin{tabular}{|l|c|c|c|c|}\hline\hline
		$L_1\leq 0.5$ &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ &  Time \\\hline \hline

1v, $?? \times 1$ & ?? & ?? & ?? & ?? \\\hline 
		3v, $?? \times 3$ & ?? & ?? & ?? & ?? \\\hline 
	 2v, $?? \times 2$ & ?? & ?? & ?? & ?? \\\hline\hline

		1v, $?? \times 1$ & ?? & ?? & ?? & ?? \\\hline 
		3v, $?? \times 3$ & ?? & ?? & ?? & ?? \\\hline 
	 2v, $?? \times 2$ & ?? & ?? & ?? & ?? \\\hline\hline

	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models 
	to obtain bounds on $\beta^{0.5}_{0,1}$ on the {\bf 20 dimension}  reduced order MNIST DNN, for timeouts of ??s, and ??s , where ??,  or 500 ($\times 1$, $\times 2$, $\times 3$) neurons use binary variables.}
\end{table}

\begin{figure*}[t!]
	\centering
\includegraphics[scale=0.5]{redimage.png} \hspace{0.8cm}
\includegraphics[scale=0.5]{redperturb.png}
\caption{An image and its perturbation from the 20-dimension reduced space with maximal $\beta^{.5}_{0,1}=0.17$ for MNIST as obtained by the "2v" $500 \times 2$ model.}
\end{figure*}

\begin{table}[h!]
	\begin{tabular}{|l|c|c|c|c|}\hline\hline
		model &    $L_1\leq 0.5$ & $L_1\leq 1$ & $L_1\leq 1.5$ &  $L_1\leq 2$ \\\hline \hline
		1v, $? \times 1$ & $80 \%$ & $32\%$ & $7\%$ & $0\%$ \\\hline
		3v, $? \times 3$ & {\bf 86 \%} & {\bf 53\%} & {\bf 20\%} & {\bf 4\%} \\\hline
		2v, $? \times 2$ & 84\% & 51\% & 16\% & 4\% \\\hline \hline
	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models 
	in the percentage of images they can certify in real-time for different values of $L_1$ perturbations.}
\end{table}

\subsection{Experiment results for regression (Pipe strain)}

	For the pipe system, to find a example of large difference on outputs is one of the main aims.

	\begin{figure*}[t!]
\includegraphics[scale=0.5]{deform.png} \hspace{0.8cm}
\includegraphics[scale=0.5]{strain.png}
\caption{2 slightly different deformations and their associated quite different strain as obtained by the "2v" $100 \times 2$ model.}
\end{figure*}

	The network of Pipe system is relatively simple and in the tests, we will open all $\ReLU$ nodes.

	We will compare the results of different modeling methods.

	\vspace*{1ex}

	\iffalse
	\begin{table}[h!]
	\begin{tabular}{|l|l|l|l|l|}\hline
		$L_1\leq 0.83$ &        Bound $\downarrow$ &  Solution $\uparrow$ &      Real $\uparrow$ &  Time \\\hline
		1v,open 100 &     {\bf 0.035613} &  0.035613 &                       0.01288 & 10608 \\\hline
		3v,open 100 &     0.040074 &  0.028934 &                      0.021441 & 10922 \\\hline
		%3v,open 100 &     0.039824 &  0.028832 &                      0.022255 & 22153 \\\hline
		2v,open 100 &     0.046719 &  0.024364 &  {\bf 0.024436} & 10922 \\\hline
	\end{tabular}
	\caption{Comparison of 1v,2v and 3v models on the pipe system with a fixed timeout of 10.000s.}
\end{table}
\fi

	\begin{table}[h!]
	\begin{tabular}{|l|c|c|c|c|}\hline\hline
		$L_1\leq 0.83$ &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ &  Time \\\hline \hline
		1v, $100 \times 1$ &     {\bf 0.0356} &  $0.0356$ & $0.0191$ &   1000s \\\hline
		3v, $100 \times 3$&     0.0414 &  0.0254 &  0.0166 &  1000s \\\hline
		2v, $100 \times 2$&     0.0418 &  0.0229 &   {\bf 0.0229} &  1000s \\\hline \hline
		3v, $90 \times 3$&      ?? &  ?? &  ?? & 14523s \\\hline
		3v, $100 \times 3$&      {\bf 0.0350} &  0.0272 &  0.0216 & 14523s \\\hline
		2v, $90 \times 2$&     ?? &  ?? &   ?? & 14523s \\\hline
		2v, $100 \times 2$&     0.0360 &  0.0236 &    {\bf 0.0236} & 14523s \\\hline \hline
		3v, $100 \times 3$&     {\bf 0.0329} &  0.0277 &  0.0165 & 72126s \\\hline
		2v, $100 \times 2$&     0.03375 &  0.0244 &  {\bf 0.0245} & 72126s \\\hline\hline
	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models on the pipe system with a fixed timeout of 1000s, 14523s and 72126s, where 90 ($\times 2$, $\times 3$) or all 100 ($\times 1, \times 2,\times 3$) neurons use binary variables.
	L1 corresponds to $3.9$ or $4$, and results should be the sum of 10 pixels, so around 10 times higher values.}
\end{table}

>>>>

	\input{a_trick}

	\input{experiments}

### 2025-07-29 AAAI2026/global.tex
global

	\section{Modeling for global verification}

@@

	We first display on Figure \ref{fig.2v} the function $f(x'_i,y_i)=\hat{y}_i$, when we know that $y_i \in [- \gamma_i, \gamma_i]$. When $x'_i < -\gamma_i$, then $x_i <0$ and we have $\hat{y}_i = \ReLU(x_i) - \ReLU(x'_i) = 0-0=0$.
    For $x'_i > \gamma_i$, we have $x_i >0$ and $\hat{y}_i=y_i$. Between these two linear planes, we have two more linear planes presenting intermediate cases, so 4 linear planes in total.

@@

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(-1,1,0)};
			\node[label={$(-\gamma_i,\gamma_i, 0)$}] at (axis cs: -1, 0.8 ,0) {};

		\end{axis}
	\end{tikzpicture}
	\caption{The outcome $\val(\hat{y}_i)=\ReLU(\val(x_i))-\ReLU(\val(x'_i))$ 
	depending on $\val(y_i) = \val(x_i)-\val(x'_i)$ and on $\val(x'_i)$.}
	\label{fig.2v}
\end{figure}

Assuming that $y_i \in [-\gamma_i, \gamma_i]$,
and that $x'_i \in [\alpha_i,\beta_i]$,
we have that $\hat{y}_i = \ReLU(x'_i+y_i)-\ReLU(x'_i)$ is the solution of:
	\begin{align*}
		& \begin{aligned}
			y_i + x'_i &\leq a\beta_i        &
			y_i &+ x'_i \geq (1-a)\alpha_i \\
			x'_i       &\leq a'\beta_i       & 
			x'_i       &\geq (1-a')\alpha_i \\
			\hat{y}_i  &\leq a\gamma_i       &
			\hat{y}_i  &\geq -a'\gamma_i \\
			\hat{y}_i  &\leq y_i + (1-a)\gamma_i  &
			\hat{y}_i  &\geq y_i - (1-a')\gamma_i \\
			\hat{y}_i  &\leq -x'_i + a\beta_i &
			\hat{y}_i  &\geq -x'_i + (1-a')\alpha_i \\
			\hat{y}_i  &\leq y_i + x'_i + (1-a)(-\alpha_i) &
			\hat{y}_i  &\geq y_i + x'_i + a'(-\beta_i)
		\end{aligned}
	\end{align*} Here, $a,a'$ are binary variables, and $a'$ is also the binary variable in the constraints for $\hat{x}'_i=\ReLU(x'_i)$.

>>>>

	\section{MILP Models for global verification}

@@

	We first display on Figure \ref{fig.2v} the function $f(x'_i,y_i)=\hat{y}_i$, when we know that $y_i \in [- \gamma_i, \gamma_i]$. When $x'_i < -\gamma_i$, then $x_i <0$ and we have $\hat{y}_i = \ReLU(x_i) - \ReLU(x'_i) = 0-0=0$.
    For $x'_i > \gamma_i$, we have $x_i >0$ and $\hat{y}_i=
    \ReLU(x_i)-\ReLU(x'_i)=x_i - x'_i = y_i$. 
    Between these two linear planes, we have two more linear planes presenting intermediate cases, so 4 linear planes in total.

@@

			\addplot3[only marks, mark=*, mark size=2pt, blue] coordinates {(-1,1,0)};
			\node[label={$(-\gamma_i,\gamma_i, 0)$}] at (axis cs: -1, 0.8 ,0) {};			
		\end{axis}
	\end{tikzpicture}
	\caption{The function computing $\hat{y}_i=\ReLU(x_i)-\ReLU(x'_i)$ 
	depending on $y_i = x_i- x'_i$ and on $x'_i$.}
	\label{fig.2v}
\end{figure}

Recall first that we use the classical MILP encoding 
for $\hat{x}'_i=\ReLU(x'_i)$, using one binary variable $a'$. We reuse 
$a'$ (with its value already settled, see Proposition \ref{Prop1}) 
in the encoding of $\hat{y}_i = \ReLU(x_i)-\ReLU(x'_i)$:

\begin{proposition}
Assuming that $y_i \in [-\gamma_i, \gamma_i]$,
and that $x'_i \in [\alpha_i,\beta_i]$,
we have that $\hat{y}_i = \ReLU(x_i=x'_i+y_i)-\ReLU(x'_i)$ is the solution of:
	\begin{align*}
		& \begin{aligned}
			y_i + x'_i &\leq a\beta_i        &
			y_i &+ x'_i \geq (1-a)\alpha_i \\
			x'_i       &\leq a'\beta_i       & 
			x'_i       &\geq (1-a')\alpha_i \\
			\hat{y}_i  &\leq a\gamma_i       &
			\hat{y}_i  &\geq -a'\gamma_i \\
			\hat{y}_i  &\leq y_i + (1-a)\gamma_i  &
			\hat{y}_i  &\geq y_i - (1-a')\gamma_i \\
			\hat{y}_i  &\leq -x'_i + a\beta_i &
			\hat{y}_i  &\geq -x'_i + (1-a')\alpha_i \\
			\hat{y}_i  &\leq y_i + x'_i + (1-a)(-\alpha_i) &
			\hat{y}_i  &\geq y_i + x'_i + a'(-\beta_i)
		\end{aligned}
	\end{align*} 
    where $a,a'$ are binary variables, 
    and $a'$ is shared with the classical MILP
    encoding for $\hat{x}'_i=\ReLU(x'_i)$.
\end{proposition}

So to encode both $\hat{y}_i$ and $\hat{x}_i$, we need 2 binary variables altogether, as the classical encoding.

%\begin{proof}
PROOF????
%\end{proof}

### 2025-07-29 AAAI2026/notation.tex
global

\begin{proposition}
\cite{MILP}
A solution $x,\hat{x},a$  of the above MILP program satisfies $\hat{x} = \ReLU(x)$,
and $a=1$ if $x> 0$ and $a=0$ if $x< 0$ (both are possible if $x=0$).
\end{proposition}

>>>>

\begin{proposition}
\cite{MILP}
\label{Prop1}
A solution $x,\hat{x},a$  of the above MILP program satisfies $\hat{x} = \ReLU(x)$,
and $a=1$ if $x> 0$ and $a=0$ if $x< 0$ (both are possible if $x=0$).
\end{proposition}
